A comparison of classification algorithms applied to the Adult Database

### 2009-12-09 report/final.tex
added preliminary description of aleph.

In terms of performance metris, we took into account the error rate and the time
spent building the rule set.

@@

>>>>

In terms of performance metrics, we took into account the error rate and the 
time spent building the rule set.

\subsection{Aleph}

Aleph \cite{aleph} is currently one of the most widely used ILP \cite{ilp} 
systems. Developed by Ashwin Srinivasan is, as of 2007, in its 5th version. The
basic Aleph algorithm (used on our experiments) consists of 4 steps:

\begin{enumerate}
    \item{Select example}
    \item{Build most-specific clause}
    \item{Search}
    \item{Remove redundant}
\end{enumerate}

Aleph uses inverse entailment to build the most-specific clause and its basic
algorithm uses a best-first search at the search step. Advanced use of Aleph 
allows tweaks in every step of the algorithm, namely:

\begin{itemize}
    \item{\textbf{Select example}: The size of the sample can be configured.}
    \item{\textbf{Build most-specific clause}: The approach on building the 
    bottom-clause can be configured. The evaluation mechanism can also be 
    tweaked.}
    \item{\textbf{Search}: A huge range of search strategies, evaluation 
    functions and refinement operators can be applied.}
    \item{\textbf{Remove redundant}: A strategy of retaining covered examples
    can be applied, for a better estimation of clause scores.}
\end{itemize}

Aleph is built entirely in prolog and is best suited to run on the YAP Prolog 
engine.

In order to use ILP to build a classifier for our data set, some data
pre-processing needs to be run, namely on converting the examples to logic 
predicates. Despite the fact that Aleph is able to deal with numerical values,
discretization has been run on the variables ``age'', ``capital\_gain'', 
``capital\_loss'' and ``hours\_per\_week''. The value distribution was the same
as described in section \ref{sec:weka}. The main classification clause is given
by \verb$more_than_50K(+person).$ Those persons winning more than 50K per year
are the positive examples, the rest are the negative examples. Each of the
attributes was converted to unary prolog predicates (for example 
\verb$young(+person).$, \verb$work_private(+person).$, \verb$male(+person).$,
\verb$is_native_of_portugal(+person).$). Missing values were simply ommited from
the background knowledge.

In terms of performance metrics, we took into account the error rate and the 
time spent building the hypothesis set. The expressiveness of the results is
also analyzed in section \ref{sec:experimental_eval}.

@@

\bibitem{ilp}
Nada Lavrac and Saso Dzeroski. Inductive Logic Programming: Techniques and Applications. Ellis Horwood, New York, 1994.
